### 2024-02-05 report/abstract/abstract_en.tex

\newpage
\pagestyle{nonumber}

\begin{center}
    \textbf{
        ABSTRACT
    }
    \\[\baselineskip]
\end{center}

\begin{center}
    \setlength\fboxsep{0.05\textwidth}\fbox{
        \parbox[t][\textwidth]{0.8\textwidth}{
            \par{
                ?
            }
        }
    }
\end{center}

\par {
    \textbf{
       Keywords (10 maximum):
    }
}

\vspace{0.3cm}
\begin{center}
    \renewcommand{\arraystretch}{1.5}\begin{tabular}{|m{0.2\textwidth}|m{0.2\textwidth}|m{0.2\textwidth}|m{0.2\textwidth}|}
        \hline
            \centering\arraybackslash{?} & \centering\arraybackslash{?} & \centering\arraybackslash{?} & \centering\arraybackslash{?} \\
        \hline
        \centering\arraybackslash{?} & \centering\arraybackslash{?} & \centering\arraybackslash{?} & \centering\arraybackslash{?} \\
        \hline
        \centering\arraybackslash{?} & \centering\arraybackslash{?} \\
        \cline{1-2}
    \end{tabular}
\end{center}