### 2024-02-14 report/test/test.tex
add list of figures

\newpage
\pagestyle{plain}

\chapter{CHAPTER TEST}

TODO
\section{SECTION TEST}

TODO

\subsection{SUBSECTION TEST}

Fig \ref{fig:test} test

\begin{figure}[h]
    \includegraphics[width=8cm]{media/epsevg_logo}
    \centering
    \caption{Test}
    \label{fig:test}
\end{figure}