### 2024-02-15 report/test/test.tex
add glossary

\newpage
\pagestyle{plain}

\chapter{CHAPTER TEST}

TODO
\section{SECTION TEST}

TODO

\subsection{SUBSECTION TEST}

Fig \ref{fig:test} test

\begin{figure}[h]
    \includegraphics[width=8cm]{media/epsevg_logo}
    \centering
    \caption{Test}
    \label{fig:test}
\end{figure}

\Glspl{matrix} are usually denoted by a bold capital letter, such as $\mathbf{A}$. The \gls{matrix}['s] $(i,j)$th element is usually denoted $a_{ij}$. \Gls{matrix} $\mathbf{I}$ is the identity \gls{matrix}. 
\gls{svm}